---
title: "FGCZ Gene Set Enrichment Analysis (GSEA)"
author: "Functional Genomics Center Zurich"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    number_sections: false
editor_options:
  chunk_output_type: console
params:
  GSEA: !r quote(fgczgseaora::GSEA)
subtitle: Using the `WebGestaltR` package
bibliography: bibliography.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

if(!exists("progress")){
  progress <- function(howmuch, detail){
    invisible(NULL)
  }
}

res <- eval(params$GSEA_res)

library(tidyverse)
library(knitr)
```

### Introduction

The following analysis compares the enrichment of particular gene/protein set members towards the upper and lower end of the provided ranked protein list (*e.g.* ranked by fold changes, *P*-values, henceforth denoted generally as *"score"*). This analysis is commonly referred to as *Gene Set Enrichment Analysis* and a more detailed description of the method can be found in @Subramanian15545. In principle, the protein list is ranked by the provided scores and an enrichment score is calculated based on the relative positions the members of a particular gene set take in the whole list. To calculate a *P*-value and a corresponding FDR, adjusted for multiplicity [@BH1995], a permutation test approach is used. The default number of permutations is $n_\text{perm}=1000$, since then the obtainable *P*-values have a lower limit of $P \geq 0.001$.

### Parameters

* Organism: `r GSEA$organism`

* Target Database: `r GSEA$target`

* Output Directory: `r GSEA$outputDir`

### GSEA Results {.tabset .tabset-pills}

#### Enriched Pathways

```{r}
GSEA$GSEA_res %>% 
  dplyr::mutate(set = paste0("[", geneSet, "](", link, ")")) %>% 
  dplyr::select(set, description, ES, NES, pValue, FDR) %>%
  knitr::kable(
    digits = 3,
    col.names = c("Gene Set", "Description", "ES", "NES", "P-value", "Q-value"),
    align = "c",
    caption = "Table of significantly enriched gene sets. ES: Enrichment Score, NES: Normalised Enrichment Score"
    )
```

#### Input Data

```{r}
colnames(GSEA$input_data) <- c("ID", "Score")
GSEA$input_data %>%
  dplyr::mutate(`Score` = round(`Score`, 2)) %>% 
  dplyr::arrange(Score) %>% 
  DT::datatable(colnames = c("ID", "Score"), width = 500, style = "bootstrap")
```

### Visualisation

```{r fig.cap="Heatmap of Score patterns in significantly enriched pathways identified by GSEA"}
GSEA$merged_data %>% 
  ggplot(aes(x = geneSet, y = geneSymbol, fill = score)) +
    geom_tile() +
    scale_fill_gradient2(low = "blue", mid = "white", high = "red") +
    labs(
      title = "Enrichment Patterns",
      subtitle = "Scores of genes from significantly enriched pathways",
      fill = "Score value"
    ) +
    xlab("Pathway") +
    ylab("Gene") +
    theme_bw()
```

```{r fig.cap="Intersection sizes between significantly enriched pathways identified by GSEA visualised using the `UpSetR` package"}
if(length(GSEA$upsetData)==1) return(message("UpSetR plot cannot be displayed. Only one pathway enriched."))

UpSetR::fromList(GSEA$upsetData) %>%
    UpSetR::upset(mb.ratio = c(0.7, 0.3))
```


### `WebGestaltR` Results

```{r results='asis'}
cat("Since `WebGestaltR` provides its own summary report, please refer to [this file](", file.path(getwd(), "../..", GSEA$outputDir) ,"/Project_GSEA_proj/Report_GSEA_proj.html)", sep = "")
```


### References


