---
title: "FGCZ Gene Set Enrichment Analysis (GSEA)"
author: "Functional Genomics Center Zurich"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    number_sections: false
editor_options:
  chunk_output_type: console
params:
  GSEA: !r quote(fgczgseaora::GSEA)
subtitle: Using the `WebGestaltR` package
bibliography: bibliography.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

if(!exists("progress")){
  progress <- function(howmuch, detail){
    invisible(NULL)
  }
}

res <- eval(params$GSEA_res)

library(tidyverse)
library(knitr)
```

### Introduction

The following analysis compares the enrichment of particular gene/protein set members towards the upper and lower end of the provided ranked protein list (*e.g.* ranked by fold changes, *P*-values, henceforth denoted generally as *"score"*). This analysis is commonly referred to as *Gene Set Enrichment Analysis* and a more detailed description of the method can be found in @Subramanian15545. In principle, the protein list is ranked by the provided scores and an enrichment score is calculated based on the relative positions the members of a particular gene set take in the whole list. To calculate a *P*-value and a corresponding FDR, adjusted for multiplicity [@BH1995], a permutation test approach is used. The default number of permutations `WebGestaltR` uses is $n_\text{perm}=1000$ \footnote{Note that due to run time issues the number of permutations is 10 per default in the `fgczgseaora` run scripts. If more accurate results are required `nperm` has to be adjusted accordingly in the run script.}.

### Parameters

* Organism: `r GSEA$organism`

* Target Database: `r GSEA$target`

* Output Directory: `r GSEA$output_dir`

* Number of permutations: `r GSEA$nperm`

### GSEA Results {.tabset .tabset-pills}

#### Enriched Pathways

```{r}
GSEA$merged_data %>% 
  distinct(geneSet, .keep_all = TRUE) %>% 
  dplyr::mutate(set = paste0("[", geneSet, "](", link, ")")) %>% 
  dplyr::select(set, description, ES, NES, pValue, FDR) %>%
  knitr::kable(
    digits = 3,
    col.names = c("Gene Set", "Description", "ES", "NES", "P-value", "Q-value"),
    align = "c",
    caption = "Table of significantly enriched gene sets. ES: Enrichment Score, NES: Normalised Enrichment Score"
    )
```

#### Input Data

```{r}
colnames(GSEA$input_data) <- c("ID", "Score")
GSEA$input_data %>%
  dplyr::mutate(`Score` = round(`Score`, 2)) %>% 
  dplyr::arrange(Score) %>% 
  DT::datatable(colnames = c("ID", "Score"), width = 500, style = "bootstrap")
```

### Visualisation

```{r fig.cap="Heatmap of Score patterns in significantly enriched pathways identified by GSEA", fig.asp=2}
GSEA$merged_data %>% 
  distinct(geneSet, .keep_all = TRUE) %>% 
  top_n(n = 10, wt = abs(NES)) %>% 
  dplyr::select(geneSet) %>% 
  inner_join(., GSEA$merged_data, by = "geneSet") -> out

out %>% 
  dplyr::select(userId, geneSet, score) %>% 
  spread(key = geneSet, value = score) -> out2
fields::image.plot(t(as.matrix(out2[,-1])), axes = FALSE)
axis(side = 1, at = seq(0, 1, length.out = ncol(out2)-1), labels = colnames(out2)[-1], lwd = 0, lwd.ticks = 0, las = 2, cex.axis = 0.8)
axis(side = 2, at = seq(0, 1, length.out = nrow(out2)), labels = out2$userId, lwd = 0, lwd.ticks = 0, las = 1, cex.axis = 0.5)
```

```{r fig.cap="Intersection sizes between significantly enriched pathways identified by GSEA visualised using the `UpSetR` package"}
toplot <- out %>%
  dplyr::select(geneSet, geneSymbol) %>%
  dplyr::mutate(IDD = 1:nrow(.)) %>%
  tidyr::spread(geneSet, geneSymbol) %>%
  dplyr::select(-IDD) %>%
  as.list() %>%
  lapply(na.omit)

if(length(GSEA$upsetData)==1) return(message("UpSetR plot cannot be displayed. Only one pathway enriched."))

UpSetR::fromList(toplot) %>%
    UpSetR::upset(mb.ratio = c(0.7, 0.3), order.by = "freq", nsets = -1)
```


### `WebGestaltR` Results

```{r results='asis'}
cat("Since `WebGestaltR` provides its own summary report, please refer to [this file](", file.path("./Project_GSEA_proj/Report_GSEA_proj.html"), ").", sep = "")
```

### References


