---
title: "Results for Webgestalt ORA analysis"
author: "FGCZ"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    includes:
      in_header: fgcz_header.html
      after_body: fgcz_footer.html
params:
  ORA: NULL
  bioinf: "Witold Wolski <wew@fgcz.ethz.ch>"
---



```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
library(purrr)
library(tidyverse)
library(flextable)

bioinf <- eval(params$bioinf)

```

The Overrepresentation analysis was run with the following settigns:

- fold change threshold `r params$fc

The fold change estimates for all contrast of all proteins are used to perform Gene Set Enrichment Analysis (GSEA) using [WebGestalt](www.webgestalt.org). 

<span style="color:red"> If some of the contrasts did NOT produce any significant gene set enrichments, they are not listed in the table below. </span>


```{r cars}

#tmp <- readRDS("D:/Dropbox/DataAnalysis/p2558_With_TOBI_K/p2558_o5358/OLD_UPLOAD/results_modelling/results_ora_24Sep2019_121117/ORA_results.Rda")
tmp <- unlist(params$ORA, recursive=FALSE)
dd <- unlist(tmp, recursive = FALSE)
ddd <- unlist(dd, recursive = FALSE)

make_gsea_path <- function(x){
  res <- list(outpath = x$outdir,
              contrast = x$fpath,
              contrast_name = x$contrast_name,
              target = x$target,
              greater = x$greater,
              threshold =  x$threshold,
              file_path = file.path(".",x$outdir, paste0("Project_", x$fname), paste0("Report_",x$fname,".html")), 
              merged_data = !is.null(x$merged_data))
  return(res)
}

paths <- purrr::map_df(ddd,make_gsea_path)

paths <- paths %>%  mutate(filter = paste0(dplyr::case_when(greater ~ "score > ", TRUE ~ "score < "  ), threshold))
paths <- paths %>% filter( merged_data == TRUE) %>% select(contrast, contrast_name, filter, target,  file_path)
#install.packages("flextable")
paths <- paths %>% dplyr::arrange(contrast, filter)
myfl <- flextable(paths) %>% set_caption("Over Representation Analysis (ORA) results using fgczgseaora and WebgestaltR. (vs - versus, gv - given).")
myfl <- myfl %>% compose( j = "file_path",
                          value = as_paragraph( hyperlink_text(x = "ora result",
                                                               url = file_path ) ) )

myfl <- myfl %>% merge_v(j = ~   contrast + contrast_name + filter)
myfl <- myfl %>% autofit() %>% theme_box() %>%  color(j = "file_path", color = "#337ab7") 
myfl
```

The ORA analysis was exectuted by : `r bioinf`.

