---
title: "Order 22513"
author: "Witold Wolski"
date: "12/02/2021"
output: html_document
params:
  resultsxx : NULL
  parametersxx: NULL
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE )
resultsV2 <- params$resultsxx
parameterV2 <- params$parametersxx
```


# Run Summary

- projectID : `r parameter$projectID`
- workunitID : `r as.character(parameter$workunitID)`
- zipfile : `r basename(parameterV2$inputMQfile)`
- clustering : `r parameterV2$clustering`
- nr.proteins : `r resultsV2$dataDims["nrPort"]`
- nr.proteins.NA.filtered : `r resultsV2$dataDims["nrPortNoNas"]`
- nr.UniprotIDs : `r resultsV2$dataDims["UniprotExtract"]`
- nr.ENTREZIDS : `r resultsV2$dataDims["ENTREZGENEID"]`
- nr.samples : `r nrow(resultsV2$prot$factors())`
- nr.of.clusters : `r resultsV2$nrCluster`
- nr.of.GS.025 : `r resultsV2$nr.of.GS.025`
- nr.of.GS.01 : `r resultsV2$nr.of.GS.01`
- median.CV : `r resultsV2$CV.50`
- median.sd : `r resultsV2$SD.50`
- id.mapping.service : `r resultsV2$id.mapping.service`


# Data QC

```{r}
prot <- resultsV2$prot
pl <- prot$get_Plotter()
pl$intensity_distribution_density()
pNa <- pl$NA_heatmap()
p <- pl$heatmap()
pc <- pl$heatmap_cor()
```

```{r namap, fig.width=10, fig.height=8, fig.cap="Heatamp based on presence and absence (black) of protein measurments"}
print(pNa)
```


```{r sampleclusth, fig.width=10, fig.height=8, fig.cap = "Heatmap with color coded protein intensities"}
print(p)
```

```{r pcaplot, fig.cap="PCA plot"}
pl$pca()

```



# Cluster proteins




```{r compareClusters, fig.cap="Cluster dendrogram."}
plot(resultsV2$dendrogram,leaflab = "none" )
```


```{r clustprofiles, fig.cap="Cluster Profiles", fig.width=10, fig.height=10}
scaledM <- resultsV2$scaledM

bb <- as_tibble(scaledM,rownames = "protein_Id")
bb <- bb %>% pivot_longer(-"protein_Id",names_to="sampleName", values_to = "transformedIntensity")
toplot <- inner_join(resultsV2$clusterAssignment, bb)
neword <- labels(as.dendrogram(hclust(dist(t(scaledM)))))
toplot$sampleName <- fct_relevel(toplot$sampleName, neword)
ggplot(toplot, aes(x=sampleName,y=transformedIntensity,group=protein_Id)) +
  geom_line(color="lightgrey") + theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~Cluster)
```


# Over Representation Analysis (ORA)

## GO Biological Process

```{r setupplot1}

figWIncrease <- 0.5
figHeightIncrease <- 0.25

P <- resultsV2$resGOEnrich$BP
figwidth <- 10
figheight <- 5
if (!is.null(P$clustProf)) {
  dt <- as.data.frame(P$clustProf)
  nrOra <- length(unique(dt$ID))
  nrClust <- length(unique(dt$Cluster))
  figwidth <- max(figwidth, 6 + nrClust*figWIncrease)
  figheight <- max(figheight, 3 + nrOra*figHeightIncrease)
}
print(c(figwidth,figheight))

```

```{r compareClustersBP, fig.height=figheight, fig.width=figwidth}
if (!is.null(P$clustProf)) {
  print(dotplot(P$clustProf,title = P$mt, showCategory	 = nrOra))
  DT::datatable(as.data.frame(P$clustProf),caption = mt)
}
```

## GO Molecular Function

```{r setupplot2}
P <- resultsV2$resGOEnrich$MF
figwidth <- 10
figheight <- 5
if (!is.null(P$clustProf)) {
  dt <- as.data.frame(P$clustProf)
  nrOra <- length(unique(dt$ID))
  nrClust <- length(unique(dt$Cluster))
  figwidth <- max(figwidth, 6 + nrClust*figWIncrease)
  figheight <- max(figheight, 3 + nrOra*figHeightIncrease)
}
```


```{r compareClustersMF,  fig.height=figheight, fig.width=figwidth}
if (!is.null(P$clustProf)) {
  print(dotplot(P$clustProf,title = P$mt, showCategory	 = nrOra))
  DT::datatable(as.data.frame(P$clustProf),caption = mt)
}
```

# GO Cellular Component

```{r setupplot3}
P <- resultsV2$resGOEnrich$CC
figwidth <- 10
figheight <- 5
if (!is.null(P$clustProf)) {
  dt <- as.data.frame(P$clustProf)
  nrOra <- length(unique(dt$ID))
  nrClust <- length(unique(dt$Cluster))
  figwidth <- max(figwidth, 6 + nrClust*figWIncrease)
  figheight <- max(figheight, 3 + nrOra*figHeightIncrease)
}

```

```{r compareClustersCC,  fig.height=figheight, fig.width=figwidth}
if (!is.null(P$clustProf)) {
  print(dotplot(P$clustProf,title = P$mt, showCategory	 = nrOra))
  DT::datatable(as.data.frame(P$clustProf),caption = mt)
}

```


